\textbf{Two-point statistics without bins: A continuous-function generalization of the correlation function estimator for large-scale structure}

### 2020-06-04 paper/paper.tex
Added section on analytic computation of RR and QQ

% there's a reason im not using k here! because these need to be vectors / tensors to then compute the amplitude
We generalize the \LS estimator defined above to any set of $K$ basis functions $f$ of the pair, so we now have
\begin{eqnarray}\displaystyle
DD &\equiv& \frac{2}{N_D(N_D-1)} \sum_{n n'} f(T_n, T_{n'}) \\
DR &\equiv& \frac{1}{N_D N_R} \sum_{n m} f(T_n, T_{m}) \\
RR &\equiv& \frac{2}{N_R(N_R-1)} \sum_{m m'} f(T_m, T_{m'}) \\
QQ &\equiv& \frac{2}{N_R(N_R-1)} \sum_{m m'} f(T_m, T_{m'}) \cdot f\T(T_m, T_{m'}),
\end{eqnarray}
%introduced this T notation to be the most general but it's a bit clunky (see below)
%what do we think about K-vector notation?
%matrix->tensor?
where $DD$, $DR$, and $RR$ are now $K$-vectors, $T$ refers to the data for the given tracer, and $QQ$ is a $K$-by-$K$ matrix defined as the outer product of basis function evaluations of the random--random pairs.

Then, we can compute the \cf as
\begin{eqnarray}\displaystyle
% here is where the T notation is weird - when you look at evaluation
a &\equiv& QQ\inv \cdot (DD + 2DR - RR) \\
\xi(T_l, T_{l'}) &\equiv& a\T \cdot f(T_l, T_{l'})
\end{eqnarray}
where $T_l$ and $T_{l'}$ contain the data values at which to evaluate $\xi$, and $a$ is a $K$-vector of the ``amplitudes'' of the basis functions.
This generalized two-point estimator, or \est, removes the need for binning as the basis functions can be continuous.
It is invariant under affine transformations; we show this in Appendix \ref{sec:affine}.

@@

>>>>

% there's a reason im not using k here! because these need to be vectors / tensors to then compute the amplitude
We generalize the \LS estimator defined above to any set of $K$ basis functions $f$ of the pair, so we now have
\begin{eqnarray}\displaystyle
DD &\equiv& \frac{2}{N_D(N_D-1)} \sum_{n n'} f(T_n, T_{n'}) \\
DR &\equiv& \frac{1}{N_D N_R} \sum_{n m} f(T_n, T_{m}) \\
RR &\equiv& \frac{2}{N_R(N_R-1)} \sum_{m m'} f(T_m, T_{m'}) \\
QQ &\equiv& \frac{2}{N_R(N_R-1)} \sum_{m m'} f(T_m, T_{m'}) \cdot f\T(T_m, T_{m'}), \label{eq:qq_proj}
\end{eqnarray}
%introduced this T notation to be the most general but it's a bit clunky (see below)
%what do we think about K-vector notation?
%matrix->tensor?
where $DD$, $DR$, and $RR$ are now $K$-vectors, $T$ refers to the data for the given tracer, and $QQ$ is a $K$-by-$K$ matrix defined as the outer product of basis function evaluations of the random--random pairs.

Then, we can compute the \cf as
\begin{eqnarray}\displaystyle
% here is where the T notation is weird - when you look at evaluation
a &\equiv& QQ\inv \cdot (DD + 2DR - RR) \\
\xi(T_l, T_{l'}) &\equiv& a\T \cdot f(T_l, T_{l'}) \label{eq:xi_proj}
\end{eqnarray}
where $T_l$ and $T_{l'}$ contain the data values at which to evaluate $\xi$, and $a$ is a $K$-vector of the ``amplitudes'' of the basis functions.
This generalized two-point estimator, or \est, removes the need for binning as the basis functions can be continuous.
It is invariant under affine transformations; we show this in Appendix \ref{sec:affine}.

@@

\section{Computing RR and QQ Analytically}\label{sec:analytic}

The autocorrelation of the random catalog is meant to approximate the window function. 
When we have a periodic cube, we can compute this $RR$ term analytically.
Here we derive this, and then derive the equivalent for our continuous-basis $RR$ and $QQ$ terms.

We consider an annulus around a single galaxy. 
This annulus has a volume $V_{ann}$.
Taking the box to have average number density $\bar{n}$, the number of galaxies expected in the annulus is $N_{ann} = A \bar{n}$, and thus this galaxy is part of $N_{ann}$ pairs.   
We do this for each of the $N_D-1$ other galaxies, and find a total of $(N_D-1) N_{annulus} = \frac{1}{2} (N_D-1) V_{ann} \bar{n}$ pairs, where the factor of $\frac{1}{2}$ accounts for the fact that this double-counts pairs.
For a cube, $\bar{n} = \frac{N_D}{L^3}$, so we finally count $\frac{1}{2} \frac{N_D}{L^3} (N-1) V_{ann}$ pairs.

For hard-edged radial bins, we can compute $A$ simply as the difference between spherical volumes. 
We can also represent this as an integral:
\begin{equation} \label{eq:vol_tophat}
V_{ann} = \int_{b_1}^{b_2} dV = 4\pi \int_{b_1}^{b_2} r^2 dr
\end{equation} 
We can generalize this to any basis function $f(r)$ that is a function of $r$:
\begin{equation}
V_{ann} = 4\pi  \int_{b_1}^{b_2} f^2(r) r^2 dr
\end{equation}
which we can see reduces to Equation \ref{eq:vol_tophat} when $f(r)$ is the tophat function (returning 1 or 0 depending on whether $r$ is between $b_1$ and $b_2$).

This gives us our full generalized analytic $RR$ term, which has elements
\begin{equation}
RR_{i,\text{ana}} = \frac{1}{2} \frac{N_D}{L^3} (N_D-1)  4\pi  \int_{0}^{r_\mathrm{max}} f_i^2(r) r^2 dr
\end{equation}
where $i$ is the index of the basis function vector.
Based on the definition of $QQ$ in Equation \ref{eq:qq_proj} as the outer product of the basis function vector and its transpose, we can see that the elements analytic $QQ$ term are:
\begin{equation}
QQ_{ij,\text{ana}} = \frac{1}{2} \frac{N_D}{L^3} (N_D-1)  4\pi  \int_{0}^{r_\mathrm{max}} f_i(r) f_j(r) r^2 dr
\end{equation}
This could be further generalized to account for basis functions that take other properties as input.

When considering a periodic box, the naive estimator is no longer biased, so we can also avoid computing the $DR$ term and calculate the amplitudes as 
\begin{equation}
a_{\mathrm{ana}} = QQ_\mathrm{ana} \inv \cdot DD.
\end{equation}
Looking back, it might have seemed strange that we use $N_D$ in calculating the analytical $RR$ term, but we now see that this normalization prefactor cancels out with that of the $DD$ term.
Finally, we can compute the correlation function as before in Equation \ref{eq:xi_proj}.
